---
title: "Homework 1"
author: "Sha Tao"
date: "September 17, 2018"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

```{r problem1_1 }

set.seed(1)

# Create a data frame
data1 = tibble(
  
  # Create a random sample of size 10 from a uniform [0, 5] distribution
  vec_unif = runif(10, min = 0, max = 5),

  # Create a logical vector indicating whether elements of the sample are greater than 2
  vec_logical = vec_unif > 2,

  # Create a (length-10) character vector
  vec_char = c("My", "name", "is", "Sha", "Tao", "and", "I", "like", "Data", "Science"),

  # Create a (length-10) factor vector
  vec_factor = factor(c("low", "low", "low", "high", "high", "low", "low", "low", "high", "high"))
)

# Print data1
data1

# Try to take the mean of the numeric variable
mean(data1$vec_unif)

# Try to take the mean of the logical variable
mean(data1$vec_logical)

# mean(data1$vec_char) doesn't work
# mean(vec_factor) doesn't work
```

The first two vectors work because a uniform [0, 5] distribution is numeric, and "true" and "false" in logical vectors are treated as 1 and 0. 
Means cannot be taken from the character vector and factor vector.

Problem 1_2 The chunk below converts variables from one type to another.

```{r problem 1_2, results = "hide"}

#applies the as.numeric function to the logical, character, and factor variables
as.numeric(data1$vec_logical)
as.numeric(data1$vec_char)
as.numeric(data1$vec_factor)

```

logical vector can be convert to numeric, and "true" and "flase"" are treated as 1 and 0.
For factor vectors, R re-orders the factors alphabetically, and assigns numeric numbers to each level. In my case, it assigns 1 to "high" and 2 to "low"."
Function as.numeric cannot treat character vector in my case and it returns 10 "NA"s. However, if I use number "1", "2" as my character vector, then it will return as numeric vector 1, 2.

Problem 1_3 The chunk below does a two-step converting.

```{r problem 1_3}

# Convert character variable from character to factor to numeric
as.numeric(as.factor(data1$vec_char))

# Convert factor variable from factor to character to numeric
as.numeric(as.character(data1$vec_factor))
```

In the first case, function as.factor converts my character vector to a factor vector. Then function as.numeric reorders the  factors alphabetically, and assigns numeric numbers to each level.
In the second case, function as.character converts my factor vector to a character vector, thus cannot be treated and returns 10 "NA"s.



